Presentation Title

### 2017-07-14 thesis/chapters/ch3-model.tex
Fix up implementation section

The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
The facilitators are elected using a \emph{luck value}, which is computed using $H(\C_r || pk_u)$,
where $\C_r$ is the consensus result 
(which can be seen as the set union of all the CP blocks collected by the facilitators)
in round $r$ and $pk_u$ is the public key of node $u$.
Intuitively, the election is guaranteed to be random 
because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.

@@

>>>>

The high communication complexity of Byzantine consensus algorithms prohibits us from running it on a large network.
Thus, for every round, we randomly select some node---called facilitators---to collect CP blocks from every other node and use them as the proposal.
The facilitators are elected using a \emph{luck value}, which is computed using $\textsf{H}(\C_r || pk_u)$,
where $\C_r$ is the consensus result 
(which can be seen as the set union of all the CP blocks collected by the facilitators)
in round $r$ and $pk_u$ is the public key of node $u$.
Intuitively, the election is guaranteed to be random 
because the output of a cryptographically secure hash function is unpredictable and $\C_r$ cannot be determined in advance.

@@

\subsection{Caching agreed fragments}

### 2017-07-14 thesis/chapters/ch5-implementation.tex
Fix up implementation section

Henceforth, we evaluate out system experimentally and compare the results with the theoretical analysis.
We begin this chapter by a description of the implementation in \Cref{sec:implementation}.
Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup}.
Finally, \Cref{sec:evaluation} presents our experimental results and our evaluation.
Our experiment primarily focuses on the consensus duration and the throughput.

@@

The prototype implementation can be found on GitHub.
\begin{displayquote}
\url{https://github.com/kc1212/consensus-thesis-code}
\end{displayquote}
It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}} and
the Twisted\footnote{\url{https://twistedmatrix.com/}} library for networking.

@@

The the goal of the experiment is to run the three protocols---consensus protocol,
transaction protocol and validation protocol---simultaneously and analyse the throughput, consensus duration and other related metrics.
We investigate these properties under the following four parameters.
\begin{enumerate}
  \item The transaction rate $r_{\text{tx}}$ per node. This is not comparable to the others because it is fixed at 2 TX/s.
        Nevertheless, it is a good value because, as we show later,
        it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
  \item The number of facilitators $n \in \{ 12, 16, \dots, 32\}$.
        Unfortunately, the maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation}.
  \item The population size $N \in \{200, \dots, 1200\}$.
        $N$ stops at 1200 is due to our physical setup, which we describe next.
  \item The two modes of transaction.
        The first mode or the ideal mode is that nodes only transact with their immediate neighbour. 
        This minimises the data volume per validated transaction because agreed fragment can be cached.
        The second mode is in the other extreme, where every transaction is with a random node out of the $N$ nodes in the system,
        thus the agreed fragment is unlikely to be cached.
\end{enumerate}

The experiment is run on the DAS-5 (The Distributed ASCI Supercomputer 5\footnote{\url{https://www.cs.vu.nl/das5/}}).
From now on, we use "machines" to refer to DAS-5 nodes and nodes to refer to a running instance in our system.
On DAS-5 we use up to 30 machine, for each machine we use 40 nodes.
This gives us the aforementioned 1200 number.
With this setup, we cannot run more nodes because the every machine only has 65535 ports available (minus the reserved ones).
But 40 nodes each need 1200 TCP connections which is 48000 TCP connections per machine and that is inching close to the limit.
While it is possible to have many more TCP connections per machine,
but it requires additional network interface which is something we do not control on the DAS-5.

@@

\section{Communication cost for a round}

The remainder of this chapter follows the same structure as the performance analysis in~\Cref{sec:performance-analysis},
where we check our theoretical evaluation with our experimental results.

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
      \caption{Transactions are with fixed neighbours}
      \label{fig:round-comms-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-communication-cost-vs-population}
      \caption{Transactions are with random neighbours}
      \label{fig:round-comms-random}
    \end{subfigure}%
  }
  \caption{}
  \label{fig:round-comms}
\end{figure}

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
      \caption{Transactions are with fixed neighbours}
      \label{fig:round-duration-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
      \caption{Transactions are with random neighbours}
      \label{fig:round-duration-random}
    \end{subfigure}%
  }
  \caption{}
  \label{fig:round-duration}
\end{figure}

\Cref{fig:round-comms} show the relationship between communication cost and population.
The most important aspect is that these results show a linear increase.
This reinforces our analytical result in~\Cref{sec:cons-complexity}.
Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost are similar.
Both peak at about 100 MB.
This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
Finally, the rate for which the communication cost increases is higher when the number of facilitators are higher.
This is also expected because the ACS algorithm has a $n^2$ term in its communication complexity.

We are also interested in how communication costs translate to time.
Hence, for the same experiment we also record the duration in seconds, which is shown on the vertical axis instead of communication cost and plot as before (\Cref{fig:consensus-duration}).
Interestingly, the duration is not entirely linear.
We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
Since if $N$ increases, every node must also perform more hash operations.
These result do not conform to analytical result in~\Cref{sec:cons-duration}.
Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
For example, the luck value can be computed by the facilitators and are sent with the consensus result.
Then the non-facilitator nodes simply checks that they are correct.

\section{Communication Cost for Transaction and Validation}

\begin{figure}[h]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
    \caption{Transactions are with fixed neighbours}
    \label{fig:tx-comms-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
    \caption{Transactions are with random neighbours}
    \label{fig:tx-comms-random}
  \end{subfigure}
  }
  \caption{Communication cost per verified transaction}
  \label{fig:tx-comms}
\end{figure}

Recall that in~\Cref{sec:communication-cost-for-tx} we argued that the communication cost per verified transaction is of $O(N)$.
To verify the argument, we plot the relationship between the communication cost and population size in~\Cref{fig:tx-comms}.
We observe a nearly linear relationship, which is due to the nearly linear relationship of the communication duration mentioned before.
Again, we believe the difference is minor and it is possible to remove the extra overhead.

@@

Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
We send validation requests at the same rate as transactions.
Upon receiving an (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
rather than just the transaction in the original validation request.
However, it may be the case that the agreed fragments arrive later than the validation request interval.
Then it is possible to have sent two or more validation requests for the transaction in the same local agreed fragment.
In this case, the remote would respond with two or more of the same agreed fragments,
which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
The result in~\Cref{fig:tx-comms-random} reinforces our argument because it is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
% It is interesting to note that some fluctuations exist in the graphs.
% This is likely due to the way how the communication cost are recorded.
% We maintain the total amount of data sent and received in a few variables.
% On every message that is sent or received, we increment the relevent variable by the size of the message.
% However, it is infeasible to log the values every time a message is sent or received as there is simply too much data for our build server to hold.
% Therefore we only log every 5 seconds.
% This does not always align with the time when a round finishes, thus some of the data may not represent reality accurately.
%  TODO this can be fixed

\section{Global Throughput}
\begin{figure}[h]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
    \caption{}
    \label{fig:global-throughput-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
    \caption{}
    \label{fig:global-throughput-random}
  \end{subfigure}
  }
  \caption{Global throughput}
  \label{fig:global-throughput}
\end{figure}

The global throughput results are shown in~\Cref{fig:global-throughput}.
Evidently, the throughput has a linear relationship with the population size.
This result is a strong indication of the horizontal scalability which we aimed to achieve.
It also matches with our analytical result.

@@

For~\Cref{fig:global-throughput-fixed},
the magnitude of our throughput may not be self-evident at first glance.
Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (4 TX/s)?
This is due to the way validated transactions are calculated.
Transactions are between two parties, hence if every node makes two transactions per seoncd,
every node also expects to receive two transactions per second.
Hence, for every node, the TX blocks are created at 4 per second.
Validation requests are sent at the same rate, which explains the magnitude.

The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is also interesting.
It is caused by the caching mechanism mentioned earlier.
If an agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
The low transaction rate in~\Cref{fig:global-throughput-random} is caused by fact that the network infrastructure cannot keep up with our demand.
We demonstrate this issue from a different perspective in~\Cref{fig:backlog}.
The graph is plotted by counting the number of transactions and the number of validated transactions every 5 seconds for one node running in a network of 1200 nodes and 32 facilitators.
In~\Cref{fig:backlog-fixed}, the number of validated transaction changes as a step function.
This means that transactions are validated in bursts, and the validation protocol can ``keep up'' with the transactions.
Note that the horizontal ``lines'' where no new validated transactions are made are roughly 70 seconds, this matches with our result in~\Cref{fig:round-duration-fixed}.
On the other hand in~\Cref{fig:backlog-random}, the validation protocol clearly cannot ``keep up'' with the rate which the transactions are made.
As a result, the global throughput is lower when transacting with random nodes than only with neighbours.

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
    \caption{}
    \label{fig:backlog-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
    \caption{}
    \label{fig:backlog-random}
  \end{subfigure}
  }
  \caption{}
  \label{fig:backlog}
\end{figure}

>>>>

To build confidence of the horizontal scalability claim from our analytical results,
we must verify the same property experimentally.
Hence, we give an open source prototype implementation of our system.
The nodes are capable of forming a network (tested up to 1500 nodes) and send transactions to each other,
which allows us to confirm our analytical claims.

We begin this chapter by a description of the implementation in \Cref{sec:implementation}.
Then, we move on to describing our experimental setup in \Cref{sec:experimental-setup} which includes a discussion on the system parameters and the physical infrastructure which we use.
Finally, \Cref{sec:comms-cost-experiment} and onwards follows the same structure as the Performance analysis (\Cref{sec:performance-analysis}),
where we analyse the performance experimentally rather than analytically.

@@

The prototype implementation can be found on GitHub.
\begin{displayquote}
\url{https://github.com/kc1212/consensus-thesis-code}
\end{displayquote}
It is written in the event driven paradigm, using the Python programming language\footnote{\url{https://www.python.org/}}.
We use the Twisted\footnote{\url{https://twistedmatrix.com/}} library for networking.

@@

The the goal of the experiment is to run the three protocols---consensus protocol,
transaction protocol and validation protocol---simultaneously and analyse the throughput, consensus duration and other related metrics.
We investigate these properties under the following four parameters.
\begin{enumerate}
  \item The transaction rate $r_{\text{tx}}$ per node. This is not comparable to the others because it is fixed at 2 TX/s.
        Nevertheless, it is a good value because, as we show later,
        it hits a bottleneck in extreme cases which helps us understand the limitations of our design.
  \item The number of facilitators $n \in \{4, 8, \dots, 32\}$.
        Unfortunately, the maximum $n$ is 32 because the limitation in liberasurecode mentioned in \Cref{sec:implementation}.
  \item The population size $N \in \{200, 300, \dots, 1200\}$.
        $N$ stops at 1200 is due to our physical setup, which we describe next.
  \item The two modes of transaction.
        The first mode is the \emph{fixed-neighbour} mode where nodes only transact with their immediate neighbour which are predetermined. 
        This minimises the data volume per validated transaction because agreed fragment can be cached.
        The second mode is in the other extreme which we call the \emph{random-neighbour} mode,
        where every transaction is with a random node out of the $N$ nodes in the system.
        It is unlikely that the agreed fragments can be reused.
\end{enumerate}

The experiment is run on the DAS-5 (The Distributed ASCI Supercomputer 5\footnote{\url{https://www.cs.vu.nl/das5/}}).
From now on, we use ``machines'' to refer to DAS-5 nodes and ``nodes'' to refer to a running instance in our system.
On DAS-5 we use up to 30 machine, for each machine we use 40 nodes.
This gives us the aforementioned 1200 number.
With this setup, we cannot run more nodes because the every machine only has 65535 ports available (minus the reserved ones).
But 40 nodes each need 1200 TCP connections which is 48000 TCP connections per machine and that is inching close to the limit.
While it is possible to have many more TCP connections per machine,
but it requires additional network interface which is something we do not control on the DAS-5.
Nevertheless, running the system with 1200 nodes gives a good indication of its scalability properties as we show later.

@@

Finally, since Bitcoin transactions are approximately 500 bytes~\cite{txsize},
we use a uniformly random transaction size sampled between 400 and 600 bytes.

\section{Communication cost for the consensus protocol}
\label{sec:comms-cost-experiment}

The remainder of this chapter follows the same structure as the performance analysis in~\Cref{sec:performance-analysis}.
We check our analytical results experimentally and again demonstrate horizontal scalability.

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-communication-cost-vs-population}
      \caption{Transactions are with fixed neighbours.}
      \label{fig:round-comms-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-communication-cost-vs-population}
      \caption{Transactions are with random neighbours.}
      \label{fig:round-comms-random}
    \end{subfigure}%
  }
  \caption{Communication cost for the consensus protocol per round increases linearly with population size.
  The error bars are larger for higher population size or higher number of facilitators is because rounds take longer thus they are repeated fewer times.}
  \label{fig:round-comms}
\end{figure}

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
    \begin{subfigure}[b]{.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-fixed/round-duration-vs-population}
      \caption{Transactions are with fixed neighbours.}
      \label{fig:round-duration-fixed}
    \end{subfigure}%
    \begin{subfigure}[b]{0.7\textwidth}
      \includegraphics[width=\textwidth]{neighbour-random/round-duration-vs-population}
      \caption{Transactions are with random neighbours.}
      \label{fig:round-duration-random}
    \end{subfigure}%
  }
  \caption{Round duration do not increase linearly with the population size. 
  This is likely due to the additional hashing operation required for larger consensus result.}
  \label{fig:round-duration}
\end{figure}

\Cref{fig:round-comms} show the relationship between communication cost of the consensus protocol per round and population size.
The most important aspect is that these results show a linear increase.
This reinforces our analytical result in~\Cref{sec:cons-complexity}.
Note that regardless of whether the transactions are performed with a random neighbour or with a fixed neighbour,
the magnitude of the communication cost are similar.
Both peak at about 100 MB.
This is expected because the consensus protocol is decoupled from the transaction protocol and the validation protocol.
Finally, the rate for which the communication cost increases is higher when the number of facilitators are higher.
This is also expected because the ACS algorithm has a $n^2$ term in its communication complexity.

We are also interested in how communication costs translate to time.
Hence, for the same experiment we also record the duration in seconds and the result is shown in~\Cref{fig:round-duration}.
Interestingly, the duration is not entirely linear.
We attribute this behaviour to the extra time needed to hash the CP blocks in the consensus result to compute the luck value.
Since if $N$ increases, every node must also perform more hash operations.
These result do not conform to analytical result in~\Cref{sec:cons-duration}.
Nevertheless, the difference is minor and there are ways to optimise the luck value computation.
For example, the luck value can be computed by the facilitators and are sent with the consensus result.
Then the non-facilitator nodes simply checks that they are correct.

\section{Communication cost for transaction and validation protocols}

\begin{figure}[h]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/tx-communication-cost-vs-population}
    \caption{Transactions are with fixed neighbours.}
    \label{fig:tx-comms-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/tx-communication-cost-vs-population}
    \caption{Transactions are with random neighbours.}
    \label{fig:tx-comms-random}
  \end{subfigure}
  }
  \caption{Communication cost per verified transaction has similar (near linear) trend as~\Cref{fig:round-duration}.
  Fluctuation for the fixed-neighbour mode exists because the cache mechanism is unpredictable.}
  \label{fig:tx-comms}
\end{figure}

We argued that the communication cost per verified transaction is of $O(N)$ (\Cref{sec:communication-cost-for-tx}).
To verify the argument, we plot the relationship between the communication cost of for every validated transaction and population size in~\Cref{fig:tx-comms}.
We observe a near linear relationship, which is due to the near linear relationship of the communication duration mentioned before.
Again, we believe the difference is minor and it is possible to remove the extra overhead.

@@

Some fluctuations exist in~\Cref{fig:tx-comms-fixed}, this is due to our caching mechanism.
We send validation requests at the same rate as transactions.
Upon receiving an (remote) agreed fragment, the caching mechanism inspects all the transactions in the agreed fragment and attempts to verify as many as it can,
rather than just the transaction in the original validation request.
However, it may be the case that the agreed fragments arrive later than the validation request interval.
Then it is possible to have sent two or more validation requests for some transactions in the same local agreed fragment.
In this case, the remote would respond with two or more of the same agreed fragments,
which results in extra (wasted) communication cost, and this is the source of the fluctuation seen in~\Cref{fig:tx-comms-fixed}.
The result in~\Cref{fig:tx-comms-random} reinforces our argument.
It is a lot more stable because for every transaction it is almost always the case that an agreed fragment is needed to validate it.
% It is interesting to note that some fluctuations exist in the graphs.
% This is likely due to the way how the communication cost are recorded.
% We maintain the total amount of data sent and received in a few variables.
% On every message that is sent or received, we increment the relevent variable by the size of the message.
% However, it is infeasible to log the values every time a message is sent or received as there is simply too much data for our build server to hold.
% Therefore we only log every 5 seconds.
% This does not always align with the time when a round finishes, thus some of the data may not represent reality accurately.
%  TODO this can be fixed

\section{Global Throughput}
\begin{figure}[h]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/throughput-vs-population}
    \caption{Transactions are with fixed neighbours.}
    \label{fig:global-throughput-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/throughput-vs-population}
    \caption{Transactions are with random neighbours.}
    \label{fig:global-throughput-random}
  \end{subfigure}
  }
  \caption{Global throughput}
  \label{fig:global-throughput}
\end{figure}

Finally, the global throughput results are shown in~\Cref{fig:global-throughput}.
Evidently, the throughput has a linear relationship with the population size.
This result is a strong indication of the horizontal scalability which we aimed to achieve.
It also matches with our analytical result.

@@

For~\Cref{fig:global-throughput-fixed},
the magnitude of our throughput may not be self-evident at first glance.
Recall that we fixed our $r_\text{tx}$ to 2, but how is it possible to have around 4800 transactions per second for 1200 nodes (which is 4 TX/s)?
This is due to the way validated transactions are calculated.
Transactions are between two parties, hence if every node makes two transactions per second,
every node also expects to receive two transactions per second.
Hence, for every node, the TX blocks are created at 4 per second.
Validation requests are sent at the same rate, which explains the magnitude.

The difference in magnitude between \Cref{fig:global-throughput-fixed} and \Cref{fig:global-throughput-random} is also interesting.
It is caused by the caching mechanism mentioned earlier.
If a new agreed fragment needs to be transmitted to validate every transaction then it puts a toll on the network infrastructure.
The low transaction rate in~\Cref{fig:global-throughput-random} is caused by fact that the network infrastructure cannot keep up with our demand.

We demonstrate the bottleneck issue from a different perspective in~\Cref{fig:backlog}.
The graph is plotted by counting the number of transactions and the number of validated transactions every 5 seconds for one node running in a network of 1200 nodes and 32 facilitators.
In~\Cref{fig:backlog-fixed}, the number of validated transaction changes as a step function.
This means that transactions are validated in bursts, and the validation protocol can ``keep up'' with the transactions.
Note that the horizontal ``lines'' where no new validated transactions are made are roughly 70 seconds, this matches with our result in~\Cref{fig:round-duration-fixed}.
On the other hand in~\Cref{fig:backlog-random}, the validation protocol clearly cannot ``keep up'' with the rate which the transactions are made.
As a result, the global throughput is lower when transacting with random nodes than only with neighbours.

\begin{figure}[tb]
  \centering
  \makebox[\linewidth][c]{%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-fixed/timeseries}
    \caption{Transactions are with fixed neighbours.}
    \label{fig:backlog-fixed}
  \end{subfigure}%
  \begin{subfigure}[b]{0.7\textwidth}
    \includegraphics[width=\textwidth]{neighbour-random/timeseries}
    \caption{Transactions are with random neighbours.}
    \label{fig:backlog-random}
  \end{subfigure}
  }
  \caption{A limitation of our system is when nodes are transacting with many random nodes,
  then the network cannot keep up with the large number of agreed fragments that need to be communicated between nodes.
  Hence the slope of validated transactions grows at a slower rate in~\Cref{fig:backlog-random}.}
  \label{fig:backlog}
\end{figure}
